% =====================================================================
%  A thermodynamically consistent framework of rate-independent
%  evolution: theory and variational approximation
%
%  Preprint version (journal-agnostic). To be reformatted with the
%  World Scientific ws-m3as class upon submission to
%  Mathematical Models and Methods in Applied Sciences (M3AS).
%
%  J. Ayensa-Jimenez, I. Romero
% =====================================================================
\documentclass[11pt,a4paper]{article}

% ---- encoding & fonts ----
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% ---- math ----
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}

% ---- layout ----
\usepackage[margin=2.6cm]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{caption}

% ---- graphics: native pgfplots (no external binaries) ----
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepgfplotslibrary{groupplots}

% ---- references ----
\usepackage[colorlinks=true,allcolors=blue!55!black]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}

% =====================================================================
%  Theorem environments
% =====================================================================
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}

% =====================================================================
%  Macros
% =====================================================================
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\Diss}{\mathrm{Diss}}
\newcommand{\Dr}{\mathcal{D}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\dd}{\mathop{}\!\mathrm{d}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\Id}{\mathrm{I}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\dist}{dist}

\crefname{assumption}{Assumption}{Assumptions}
\Crefname{assumption}{Assumption}{Assumptions}

% =====================================================================
\begin{document}

\title{\bfseries Rate-Independent Epigenetics: a thermodynamically
consistent framework for modelling epigenetic changes}

\author{Jacobo Ayensa-Jim\'enez\thanks{Technical University of Madrid
(UPM), Madrid, Spain. \texttt{jacobo.ayensa@upm.es}.}
\and
Ignacio Romero\thanks{Technical University of Madrid (UPM) and IMDEA
Materials Institute, Madrid, Spain. \texttt{ignacio.romero@upm.es}.}}

\date{\today}

\maketitle

\begin{abstract}
Epigenetic changes---heritable, long-lived, yet actively reversible
modifications of the chromatin state---display memory, threshold
activation and hysteresis, features that are the hallmark of
\emph{rate-independent} dissipative evolution. We propose a
mathematical framework, which we call Rate-Independent Epigenetics
(RIE), that models epigenetic change within the theory of
rate-independent systems and is, by construction, consistent with the
laws of thermodynamics. A model is specified by a triple: a state
space of epigenetic configurations, a stored-energy functional
depending on the state and on an external (micro-environmental)
loading, and a $1$-homogeneous dissipation potential encoding the
resistance of the epigenetic machinery to change. Assuming an
energetic evolution principle, the governing equations---an energetic
(global stability plus energy balance) formulation and an equivalent
subdifferential inclusion of Biot type---follow systematically, with
no further modelling hypotheses. Solutions satisfy, a priori, energetic
estimates that coincide with the first and second laws: the energy
balance is exact energy conservation, and the $1$-homogeneity of the
dissipation potential forces a non-negative, minimal (\emph{economical})
dissipation. Under natural coercivity and continuity assumptions we
establish existence of energetic solutions and, via vanishing
viscosity, of balanced-viscosity solutions that resolve the ambiguity
of the energetic formulation at epigenetic switches; uniqueness holds
under convexity and, in the scalar case, under a mild
finite-multiplicity condition. We then build a variational time
integrator, prove its convergence to energetic solutions and a global
energy-consistency estimate, and reduce it, in the scalar case, to a
two-sided return map. The framework is illustrated on the scalar linear
play operator, the minimal RIE model of epigenetic memory, for which
the state evolution is obtained in closed form and reproduces both
reversible and irreversible (locked-in) responses; the integrator
recovers the exact solution at the nodes and exhibits the predicted
first-order energy consistency. The bistable model of catastrophic
phenotype switching, and its biological calibration, are developed in a
companion paper.

\medskip
\noindent\textbf{Keywords:} rate-independent systems; epigenetics;
energetic solutions; balanced-viscosity solutions; dissipation
potential; hysteresis; play operator; variational time integration;
thermodynamic consistency.

\medskip
\noindent\textbf{AMS Subject Classification:}
49J40, 74C05, 34G25, 74N30, 65M12, 92C40, 80A17.
\end{abstract}

\tableofcontents
\bigskip

\input{sections/introduction}
\input{sections/framework}
\input{sections/thermodynamics}
\input{sections/mathematical-results}
\input{sections/numerical-methods}
\input{sections/experiments}
\input{sections/conclusions}

\appendix
\input{sections/appendix}

\bibliographystyle{plain}
\bibliography{references}

\end{document}
