% =====================================================================
%  Appendices — supporting results and detailed proofs
% =====================================================================
%
%  The appendices are organised so that each result depends only on
%  results stated before it: the instrumental lemmas come first and no
%  proof appeals to a later statement. Results already stated in the
%  body (the Biot inclusion and the main existence, uniqueness,
%  convergence and consistency theorems) are recalled in italics at the
%  head of their proofs, so the reader need not turn back for the
%  hypotheses.
%
%  Two classical facts are used repeatedly and without further comment.
%  By \emph{Rademacher's theorem}, a Lipschitz function is
%  differentiable almost everywhere; more generally, an absolutely
%  continuous $u:[0,T]\to\R^n$ is differentiable a.e., its derivative
%  $\dot u$ is Lebesgue integrable, and the fundamental theorem of
%  calculus holds, $u(t)-u(s)=\int_s^t\dot u$. We write $\dot u$ for this
%  a.e.\ derivative throughout.

\section{Preliminaries and the governing equation}
\label{app:framework}

Throughout this appendix we use, without further comment, two classical
facts. By \emph{Rademacher's theorem} a Lipschitz function
$u:[0,T]\to\R^n$ is differentiable almost everywhere; and, more
generally, an absolutely continuous $u$ is differentiable a.e., its
derivative $\dot u$ is Lebesgue integrable, and the fundamental theorem
of calculus for the Lebesgue integral holds,
$u(t)-u(s)=\int_s^t\dot u(\eta) \, \dd \eta$. This justifies writing $\dot u$ for the
a.e.\ derivative of every absolutely continuous or Lipschitz function
below.

\subsection{Convex-analytic and metric preliminaries}

\begin{lemma}[Saturation identity and scale invariance]\label{lem:saturation}
Let $\Psi\colon\R^n\to[0,+\infty)$ be convex, normalised
($\Psi(\bs 0)=0$) and positively $1$-homogeneous, and recall
$\E:=\partial\Psi(\bs 0)=\{\bs f:\bs f\cdot\bs w\le\Psi(\bs w)\ \forall\,\bs w\in\R^n\}$.
Then:
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=2.4em,itemsep=2pt]
  \item\label{sat:i} for every $\bs v\in\R^n$ and every
  $\bs f\in\partial\Psi(\bs v)$,
  \begin{equation}\label{eq:saturation}
    \bs f\cdot\bs v=\Psi(\bs v);
  \end{equation}
  \item\label{sat:ii} $\bs f\in\partial\Psi(\bs v)$ if and only if
  $\bs f\in\E$ \emph{and} $\bs f\cdot\bs v=\Psi(\bs v)$;
  \item\label{sat:iii} $\partial\Psi(\lambda\bs v)=\partial\Psi(\bs v)$
  for every $\lambda>0$.
\end{enumerate}
\end{lemma}

\begin{proof}
\emph{\ref{sat:i}.} By definition of the convex subdifferential,
$\bs f\in\partial\Psi(\bs v)$ means
\begin{equation}\label{eq:subdiff_def}
  \Psi(\bs w)\ge\Psi(\bs v)+\bs f\cdot(\bs w-\bs v)
  \qquad\forall\,\bs w\in\R^n .
\end{equation}
Take $\bs w=\lambda\bs v$ with $\lambda>0$. Positive $1$-homogeneity
gives $\Psi(\lambda\bs v)=\lambda\Psi(\bs v)$, so~\eqref{eq:subdiff_def}
becomes
\[
  \lambda\Psi(\bs v)\ge\Psi(\bs v)+(\lambda-1)\,\bs f\cdot\bs v,
  \qquad\text{i.e.}\qquad
  (\lambda-1)\bigl(\Psi(\bs v)-\bs f\cdot\bs v\bigr)\ge0 .
\]
For $\lambda>1$ the factor $(\lambda-1)$ is positive, whence
$\Psi(\bs v)\ge\bs f\cdot\bs v$; for $\lambda\in(0,1)$ it is negative,
whence $\Psi(\bs v)\le\bs f\cdot\bs v$. Combining the two
gives~\eqref{eq:saturation}.

\emph{\ref{sat:ii}.} ($\Rightarrow$) If $\bs f\in\partial\Psi(\bs v)$,
then~\eqref{eq:saturation} holds, and substituting
$\bs f\cdot\bs v=\Psi(\bs v)$ into the right-hand side
of~\eqref{eq:subdiff_def} yields $\Psi(\bs w)\ge\bs f\cdot\bs w$ for all
$\bs w$, i.e.\ $\bs f\in\E$. ($\Leftarrow$) Conversely, if $\bs f\in\E$
(so $\bs f\cdot\bs w\le\Psi(\bs w)$ for all $\bs w$) and
$\bs f\cdot\bs v=\Psi(\bs v)$, then for every $\bs w$
\[
  \Psi(\bs v)+\bs f\cdot(\bs w-\bs v)
  =\bs f\cdot\bs v+\bs f\cdot(\bs w-\bs v)
  =\bs f\cdot\bs w\le\Psi(\bs w),
\]
which is exactly~\eqref{eq:subdiff_def}; hence
$\bs f\in\partial\Psi(\bs v)$.

\emph{\ref{sat:iii}.} Let $\lambda>0$. By~\ref{sat:ii},
$\bs f\in\partial\Psi(\bs v)$ iff $\bs f\in\E$ and
$\bs f\cdot\bs v=\Psi(\bs v)$. Multiplying the scalar identity by
$\lambda$ and using $1$-homogeneity,
$\bs f\cdot(\lambda\bs v)=\lambda\,\bs f\cdot\bs v=\lambda\Psi(\bs v)=\Psi(\lambda\bs v)$;
since the membership $\bs f\in\E$ does not involve $\bs v$, applying
\ref{sat:ii} again (now at the point $\lambda\bs v$) gives
$\bs f\in\partial\Psi(\lambda\bs v)$. The two conditions are symmetric
in $\bs v$ and $\lambda\bs v$, so the subdifferentials coincide.
\end{proof}

\begin{proposition}[Linearity and continuity of the dissipation
distance]\label{prop:Dr_linear}
If $\Q$ is convex and $\Psi$ is continuous, convex and positively
$1$-homogeneous, then
\begin{equation}\label{eq:Dr_linear}
  \Dr(\bs q_1,\bs q_2)=\Psi(\bs q_2-\bs q_1)\qquad\forall\,\bs q_1,\bs q_2\in\Q,
\end{equation}
and $\Dr$ is continuous on $\Q\times\Q$. In particular
$\Dr(q_1,q_2)=\rho|q_2-q_1|$ when $\Psi(\dot q)=\rho|\dot q|$.
\end{proposition}

\begin{proof}
\emph{Lower bound.} Let $\gamma\in\mathrm{AC}([0,1];\Q)$ join
$\bs q_1$ to $\bs q_2$. Jensen's inequality for the convex function
$\Psi$ applied to the (probability) average of $\dot\gamma$ over
$[0,1]$ gives
\[
  \int_0^1\Psi(\dot\gamma(s))\dd s
  \ge\Psi\!\Bigl(\int_0^1\dot\gamma(s)\dd s\Bigr)
  =\Psi(\gamma(1)-\gamma(0))=\Psi(\bs q_2-\bs q_1).
\]
Taking the infimum over $\gamma$ yields
$\Dr(\bs q_1,\bs q_2)\ge\Psi(\bs q_2-\bs q_1)$.

\emph{Upper bound.} Since $\Q$ is convex, the straight segment
$\bar\gamma(s)=\bs q_1+s(\bs q_2-\bs q_1)$ lies in $\Q$ and is
admissible, with constant velocity $\dot{\bar\gamma}\equiv\bs q_2-\bs q_1$
and cost $\int_0^1\Psi(\bs q_2-\bs q_1)\dd s=\Psi(\bs q_2-\bs q_1)$.
Hence $\Dr(\bs q_1,\bs q_2)\le\Psi(\bs q_2-\bs q_1)$, and~\eqref{eq:Dr_linear}
follows. Continuity of $\Dr$ is then immediate from the continuity of
$\Psi$ and of $(\bs q_1,\bs q_2)\mapsto\bs q_2-\bs q_1$.
\end{proof}

\begin{proposition}[Metric properties of the dissipation
distance]\label{prop:quasi_distance}
Under \ref{A:Psi} and $\Q$ convex, the dissipation distance satisfies,
for all $\bs q_1,\bs q_2,\bs q_3\in\Q$:
\textnormal{(D1)} $\Dr(\bs q_1,\bs q_2)\ge0$ and $\Dr(\bs q,\bs q)=0$;
\textnormal{(D2)} $\Dr(\bs q_1,\bs q_3)\le\Dr(\bs q_1,\bs q_2)+\Dr(\bs q_2,\bs q_3)$.
Since $\Psi$ is symmetric, $\Dr$ is a pseudodistance; if in addition
$\Psi(\bs v)=0\Rightarrow\bs v=\bs 0$ (in particular under
\ref{A:coerc}), $\Dr$ is a distance.
\end{proposition}

\begin{proof}
By \cref{prop:Dr_linear}, $\Dr(\bs q_1,\bs q_2)=\Psi(\bs q_2-\bs q_1)$;
the four properties are then inherited from those of $\Psi$.
\emph{(D1).} $\Dr(\bs q_1,\bs q_2)=\Psi(\bs q_2-\bs q_1)\ge0$ by
positiveness, and $\Dr(\bs q,\bs q)=\Psi(\bs 0)=0$ by normalisation.
\emph{(D2).} A convex, positively $1$-homogeneous $\Psi$ is subadditive:
$\Psi(\bs a+\bs b)=2\,\Psi\bigl(\tfrac12\bs a+\tfrac12\bs b\bigr)
\le2\bigl(\tfrac12\Psi(\bs a)+\tfrac12\Psi(\bs b)\bigr)=\Psi(\bs a)+\Psi(\bs b)$.
Applying this to $\bs a=\bs q_2-\bs q_1$ and $\bs b=\bs q_3-\bs q_2$,
\[
  \Dr(\bs q_1,\bs q_3)=\Psi(\bs q_3-\bs q_1)
  =\Psi\bigl((\bs q_2-\bs q_1)+(\bs q_3-\bs q_2)\bigr)
  \le\Psi(\bs q_2-\bs q_1)+\Psi(\bs q_3-\bs q_2)
  =\Dr(\bs q_1,\bs q_2)+\Dr(\bs q_2,\bs q_3).
\]
\emph{Symmetry and definiteness.} If $\Psi(-\bs v)=\Psi(\bs v)$ then
$\Dr(\bs q_2,\bs q_1)=\Psi(\bs q_1-\bs q_2)=\Psi(\bs q_2-\bs q_1)=\Dr(\bs q_1,\bs q_2)$,
a pseudodistance; and if $\Psi(\bs v)=0\Rightarrow\bs v=\bs 0$ then
$\Dr(\bs q_1,\bs q_2)=0$ forces $\bs q_2=\bs q_1$, so $\Dr$ separates
points.
\end{proof}

\begin{lemma}[Cumulated dissipation of an absolutely continuous
path]\label{lem:diss_ac}
Let $\bs q\in\mathrm{AC}([0,T];\Q)$ with $\Q$ convex. Then
\begin{equation}\label{eq:diss_ac}
  \Diss_\Psi(\bs q;[0,t])=\int_0^t\Psi(\dot{\bs q}(s))\dd s
  \qquad\text{for every }t\in[0,T],
\end{equation}
so $t\mapsto\Diss_\Psi(\bs q;[0,t])$ is absolutely continuous with
$\tfrac{\dd}{\dd t}\Diss_\Psi(\bs q;[0,t])=\Psi(\dot{\bs q}(t))$ for a.e.\
$t$.
\end{lemma}

\begin{proof}
For a partition $0=t_0<\dots<t_N=t$, \cref{prop:Dr_linear} and the
fundamental theorem of calculus give
\[
  \sum_{i=1}^{N}\Dr(\bs q(t_{i-1}),\bs q(t_i))
  =\sum_{i=1}^{N}\Psi\bigl(\bs q(t_i)-\bs q(t_{i-1})\bigr)
  =\sum_{i=1}^{N}\Psi\!\Bigl(\int_{t_{i-1}}^{t_i}\dot{\bs q}(s)  \dd s \Bigr).
\]
Taking the supremum over all partitions, the left side is
$\Diss_\Psi(\bs q;[0,t])$ by definition, while the supremum of the right
side is $\int_0^t\Psi(\dot{\bs q}(s))\dd s$: the inequality "$\le$" follows from
the subadditivity of $\Psi$ (as in \cref{prop:quasi_distance}) together
with Jensen's inequality on each subinterval, and "$\ge$" from
refining the partition, this being precisely the definition of the
$\Psi$-variation of an absolutely continuous curve. This
proves~\eqref{eq:diss_ac}; absolute continuity of $t\mapsto\int_0^t\Psi(\dot{\bs q}(s))\dd s$
and the stated derivative are then standard, since
$s\mapsto\Psi(\dot{\bs q}(s))\in L^1(0,T)$ (as $\Psi$ is Lipschitz on
bounded sets and $\dot{\bs q}\in L^1$).
\end{proof}

\subsection{Energy estimates}

\begin{proposition}[Coercivity, absolute continuity and power
control]\label{prop:power}
Under \ref{A:W}--\ref{A:S}, the following hold.
\begin{enumerate}[label=\textnormal{(E\arabic*)},leftmargin=2.6em,itemsep=2pt]
  \item\label{E:compact} For every $c\in\R$ and $S\ge0$, the sublevel
  set $\mathcal L_c(S):=\{\bs q\in\Q:E(\bs q,S)\le c\}$ is compact.
  \item\label{E:AC} For each fixed $\bs q\in\Q$, $t\mapsto E(\bs q,S(t))$
  is absolutely continuous on $[0,T]$, with
  $\frac{\dd}{\dd t}E(\bs q,S(t))=-\bs q\cdot\bs\ell'(S(t))\dot S(t)$ for
  a.e.\ $t$.
  \item\label{E:gronwall} There exist constants $c_3,c_4>0$, depending
  only on $c_1,c_2,p,C_\ell,C_{\ell'}$, such that
  $|\partial_S E(\bs q,S)|\le c_3\bigl(E(\bs q,S)+c_4\bigr)$ for all
  $\bs q\in\Q$, $S\ge0$; moreover $E(\bs q,S)+c_4\ge1>0$.
\end{enumerate}
\end{proposition}

\begin{proof}
\emph{\ref{E:compact}.} \textit{Boundedness.} Let
$\bs q\in\mathcal L_c(S)$, i.e.\ $W(\bs q)-\bs q\cdot\bs\ell(S)\le c$.
By \ref{A:W}, \ref{A:ell} and Cauchy--Schwarz,
\begin{equation}\label{eq:coercive_sublevel}
  c_1\|\bs q\|^p-c_2\le W(\bs q)\le c+\bs q\cdot\bs\ell(S)\le c+C_\ell\|\bs q\|,
\end{equation}
so $c_1\|\bs q\|^p-C_\ell\|\bs q\|\le c+c_2$. Since $p>1$, the
left-hand side tends to $+\infty$ as $\|\bs q\|\to\infty$; therefore
$\|\bs q\|$ is bounded uniformly on $\mathcal L_c(S)$. \textit{Closedness.}
If $\bs q_n\in\mathcal L_c(S)$ and $\bs q_n\to\bs q$ in $\Q$ (closed),
continuity of $W$ (\ref{A:W}) and of $\bs\ell$ (\ref{A:ell}, Lipschitz)
give $E(\bs q,S)=\lim_n E(\bs q_n,S)\le c$, so $\bs q\in\mathcal L_c(S)$.
A closed, bounded subset of $\R^n$ is compact by Heine--Borel.

\emph{\ref{E:AC}.} By \ref{A:ell}, $\bs\ell$ is Lipschitz, hence, by
Rademacher's theorem, differentiable a.e.\ with $\|\bs\ell'\|\le C_{\ell'}$;
by \ref{A:S}, $S\in W^{1,1}(0,T)\subset\mathrm{AC}([0,T])$. The
composition $t\mapsto\bs\ell(S(t))$ is therefore absolutely continuous,
with $\frac{\dd}{\dd t}\bs\ell(S(t))=\bs\ell'(S(t))\dot S(t)$ a.e.\ by the
chain rule for Sobolev functions. As $W(\bs q)$ does not depend on $t$,
$t\mapsto E(\bs q,S(t))=W(\bs q)-\bs q\cdot\bs\ell(S(t))$ is absolutely
continuous with the stated derivative; its integrability follows from
$|\bs q\cdot\bs\ell'(S(t))\dot S(t)|\le C_{\ell'}\|\bs q\|\,|\dot S(t)|$
and $\dot S\in L^1$.

\emph{\ref{E:gronwall}.} Since $|\partial_S E(\bs q,S)|=|\bs q\cdot\bs\ell'(S)|\le C_{\ell'}\|\bs q\|$,
it suffices to bound $\|\bs q\|$ by $E(\bs q,S)$. From
\eqref{eq:coercive_sublevel} with $c=E(\bs q,S)$,
\begin{equation}\label{eq:coercive_rearranged}
  c_1\|\bs q\|^p-C_\ell\|\bs q\|\le E(\bs q,S)+c_2 .
\end{equation}
Apply Young's inequality $ab\le a^p/p+b^{p'}/p'$ (with
$\tfrac1p+\tfrac1{p'}=1$) to
$a=(c_1p/2)^{1/p}\|\bs q\|$ and $b=C_\ell/(c_1p/2)^{1/p}$, so that
$ab=C_\ell\|\bs q\|$:
\begin{equation}\label{eq:young}
  C_\ell\|\bs q\|\le\frac{c_1}{2}\|\bs q\|^p+K,
  \qquad K:=\frac{1}{p'}\Bigl(\frac{C_\ell}{(c_1p/2)^{1/p}}\Bigr)^{p'}>0 .
\end{equation}
Substituting~\eqref{eq:young} into~\eqref{eq:coercive_rearranged} and
absorbing,
$\tfrac{c_1}{2}\|\bs q\|^p\le E(\bs q,S)+c_2+K=:E(\bs q,S)+\tilde c_2$;
in particular $E(\bs q,S)+\tilde c_2\ge\tfrac{c_1}{2}\|\bs q\|^p\ge0$,
and
$\|\bs q\|\le(2/c_1)^{1/p}\bigl(E(\bs q,S)+\tilde c_2\bigr)^{1/p}$.
Since $1/p<1$, $r^{1/p}\le r+1$ for all $r\ge0$; applying this to
$r=E(\bs q,S)+\tilde c_2$ gives
\[
  |\partial_S E(\bs q,S)|\le C_{\ell'}\|\bs q\|
  \le C_{\ell'}\Bigl(\tfrac{2}{c_1}\Bigr)^{1/p}\bigl(E(\bs q,S)+\tilde c_2+1\bigr)
  = c_3\bigl(E(\bs q,S)+c_4\bigr),
\]
with $c_3=C_{\ell'}(2/c_1)^{1/p}$ and $c_4=\tilde c_2+1$; and
$E(\bs q,S)+c_4=(E(\bs q,S)+\tilde c_2)+1\ge1$.
\end{proof}

\subsection{The Biot inclusion}

\begin{proof}[Proof of \cref{prop:biot}]
\emph{Statement.} \emph{Let $\bs q\in\mathrm{AC}([0,T];\Q)$ satisfy
\textnormal{(S)} and~\textnormal{(E)} with $E(\cdot,S)$ differentiable.
Then $\bs f(\bs q,S)\in\partial\Psi(\dot{\bs q})$ for a.e.\ $t$
$($equivalently~\eqref{eq:biot}$)$; conversely, if $E(\cdot,S)$ is
convex, \eqref{eq:biot} together with $\bs f(\bs q,S)\in\E$ implies
\textnormal{(S)}--\textnormal{(E)}.}

\smallskip
Let $t$ be a point at which $\dot{\bs q}(t)$ exists (a.e.\ $t$, since
$\bs q\in\mathrm{AC}$) and at which the energy balance~(E) is
differentiable.

\emph{Step 1: local stability.} By the global stability~(S), the state
$\bs q(t)$ minimises $\bs r\mapsto E(\bs r,S(t))+\Dr(\bs q(t),\bs r)$
over $\Q$; in particular it is a local minimiser. Since
$E(\cdot,S(t))$ is differentiable and, by \cref{prop:Dr_linear},
$\Dr(\bs q(t),\bs r)=\Psi(\bs r-\bs q(t))$ has subdifferential
$\partial\Psi(\bs r-\bs q(t))$, which at $\bs r=\bs q(t)$ equals
$\partial\Psi(\bs 0)=\E$, the first-order optimality condition reads
$\bs 0\in\partial_{\bs q}E(\bs q(t),S(t))+\E$, that is,
\begin{equation}\label{eq:local_stab}
  \bs f(\bs q(t),S(t))=-\partial_{\bs q}E(\bs q(t),S(t))\in\E .
\end{equation}

\emph{Step 2: energy identity along the trajectory.} By
\cref{prop:power}\ref{E:AC} the map $t\mapsto E(\bs q(t),S(t))$ is
absolutely continuous, and by \cref{lem:diss_ac} so is
$t\mapsto\Diss_\Psi(\bs q;[0,t])$, with derivative $\Psi(\dot{\bs q})$.
Differentiating the balance~(E),
$E(\bs q(t),S(t))+\Diss_\Psi(\bs q;[0,t])=E(\bs q(0),S(0))+\int_0^t\partial_S E\,\dot S$,
at a.e.\ $t$ gives
\[
  \partial_{\bs q}E(\bs q,S)\cdot\dot{\bs q}+\partial_S E(\bs q,S)\,\dot S
  +\Psi(\dot{\bs q})=\partial_S E(\bs q,S)\,\dot S,
\]
and the terms $\partial_S E\,\dot S$ cancel, leaving
\begin{equation}\label{eq:power_identity}
  \bs f(\bs q,S)\cdot\dot{\bs q}
  =-\partial_{\bs q}E(\bs q,S)\cdot\dot{\bs q}
  =\Psi(\dot{\bs q}).
\end{equation}

\emph{Step 3: conclusion.} By~\eqref{eq:local_stab},
$\bs f\in\E$; by~\eqref{eq:power_identity},
$\bs f\cdot\dot{\bs q}=\Psi(\dot{\bs q})$. \Cref{lem:saturation}\ref{sat:ii}
(with $\bs v=\dot{\bs q}$) then gives
$\bs f\in\partial\Psi(\dot{\bs q})$, i.e.\ the Biot
inclusion~\eqref{eq:biot}.

\emph{Converse.} Assume $E(\cdot,S)$ convex and that both~\eqref{eq:biot}
and $\bs f(\bs q,S)\in\E$ hold a.e. Convexity makes local and global
stability equivalent, so $\bs f\in\E$ is~(S). For~(E): by
\cref{lem:saturation}\ref{sat:i}, \eqref{eq:biot} gives
$\bs f\cdot\dot{\bs q}=\Psi(\dot{\bs q})$ a.e.; integrating and using
\cref{lem:diss_ac} together with \cref{prop:power}\ref{E:AC} reverses
Step~2 and recovers the balance~(E).
\end{proof}

% =====================================================================
\section{Well-posedness}
\label{app:math}

\subsection{Compatibility}

\begin{lemma}[Compatibility conditions]\label{lem:compatibility}
Under \ref{A:W}--\ref{A:Psi}, let $(t_m,\bs q_m)_{m\in\N}$ be a
\emph{stable sequence}, i.e.\ $\bs q_m\in\mathcal S(t_m)$ and
$\sup_m E(\bs q_m,S(t_m))<\infty$, with $t_m\to t$ in $[0,T]$ and
$\bs q_m\to\bs q$ in $\Q$. Then:
\begin{enumerate}[label=\textnormal{(C\arabic*)},leftmargin=2.6em,itemsep=2pt]
  \item\label{C1} for a.e.\ $t$, $\partial_S E(\bs q_m,S(t_m))\to\partial_S E(\bs q,S(t))$;
  \item\label{C2} $\bs q\in\mathcal S(t)$.
\end{enumerate}
\end{lemma}

\begin{proof}
\emph{\ref{C1}.} At any $t\notin\mathcal N$ (the null set outside which
$\dot S$ exists), $\partial_S E(\bs q,S(t))=-\bs q\cdot\bs\ell'(S(t))\dot S(t)$
is linear and continuous in $\bs q$. Since $S\in W^{1,1}(0,T)\subset C([0,T])$
and $\bs\ell'$ is bounded and (a.e.) continuous, $t_m\to t$ gives
$\bs\ell'(S(t_m))\dot S(t_m)\to\bs\ell'(S(t))\dot S(t)$; combined with
$\bs q_m\to\bs q$,
\[
  \partial_S E(\bs q_m,S(t_m))=-\bs q_m\cdot\bs\ell'(S(t_m))\dot S(t_m)
  \longrightarrow-\bs q\cdot\bs\ell'(S(t))\dot S(t)=\partial_S E(\bs q,S(t)).
\]
\emph{\ref{C2}.} By $\bs q_m\in\mathcal S(t_m)$, for every fixed
$\tilde{\bs q}\in\Q$,
\begin{equation}\label{eq:stability_m}
  E(\bs q_m,S(t_m))\le E(\tilde{\bs q},S(t_m))+\Dr(\bs q_m,\tilde{\bs q}).
\end{equation}
As $m\to\infty$: the left side tends to $E(\bs q,S(t))$ by continuity of
$W$ and $\bs\ell$ and $S(t_m)\to S(t)$; the first right-hand term tends
to $E(\tilde{\bs q},S(t))$ for the same reason; the second tends to
$\Dr(\bs q,\tilde{\bs q})$ by continuity of $\Dr$
(\cref{prop:Dr_linear}). Passing to the limit in~\eqref{eq:stability_m},
$E(\bs q,S(t))\le E(\tilde{\bs q},S(t))+\Dr(\bs q,\tilde{\bs q})$ for
every $\tilde{\bs q}$, i.e.\ $\bs q\in\mathcal S(t)$.
\end{proof}

\subsection{Existence and uniqueness of energetic solutions}

\begin{proof}[Proof of \cref{thm:existence}]
\emph{Statement.} \emph{Under \ref{A:W}--\ref{A:Psi}, for every
$\bs q_0\in\mathcal S(0)$ there exists at least one energetic solution
$\bs q:[0,T]\to\Q$ of bounded variation.}

\smallskip
We verify the hypotheses of the abstract existence theorem
\cite[Thm.~2.1.6]{MielkeRoubicek} for the rate-independent system
$(\Q,E,\Dr)$.
\begin{itemize}[leftmargin=1.4em,itemsep=2pt]
  \item \emph{Dissipation distance.} By \cref{prop:quasi_distance},
  $\Dr$ satisfies (D1)--(D2); since $\Psi$ is symmetric (\ref{A:Psi}),
  $\Dr$ is a genuine (extended) quasidistance, and under \ref{A:coerc}
  a distance --- in all cases enough for the cited theorem.
  \item \emph{Compactness of sublevels.} \ref{A:W} and \ref{A:ell} are
  the hypotheses of \cref{prop:power}\ref{E:compact}, which provides
  the required compactness of $\mathcal L_c(S)$ for all $c,S$.
  \item \emph{Control of the power.} \cref{prop:power}\ref{E:AC}
  gives the absolute continuity of $t\mapsto E(\bs q,S(t))$, and
  \ref{E:gronwall} the bound $|\partial_S E(\bs q,S)|\le c_3(E(\bs q,S)+c_4)$.
  Together, $\bigl|\tfrac{\dd}{\dd t}E(\bs q,S(t))\bigr|\le c_3\bigl(E(\bs q,S(t))+c_4\bigr)|\dot S(t)|$,
  i.e.\ the abstract power-control condition holds with
  $\alpha_E(t)=c_3|\dot S(t)|\in L^1(0,T)$ by \ref{A:S}.
  \item \emph{Compatibility.} \cref{lem:compatibility} supplies (C1)
  (continuity of the reduced power along stable sequences) and (C2)
  (closedness of the stability set), the two compatibility conditions
  required.
  \item \emph{Topology.} $\Q\subset\R^n$ with the Euclidean topology is
  separable and metrizable.
\end{itemize}
All hypotheses being met and $\bs q_0\in\mathcal S(0)$ by assumption,
\cite[Thm.~2.1.6]{MielkeRoubicek} yields an energetic solution
$\bs q:[0,T]\to\Q$ of bounded variation in the sense of
\cref{def:energetic}.
\end{proof}

\begin{proof}[Proof of \cref{thm:uniqueness}]
\emph{Statement.} \emph{Under \ref{A:W}--\ref{A:Psi}, if $W\in C^3(\Q)$
is uniformly convex $(D^2W\succeq\alpha\Id,\ \alpha>0)$ and
$\bs\ell\in C^3$, $S\in C^3([0,T])$, then for every
$\bs q_0\in\mathcal S(0)$ the energetic solution is unique.}

\smallskip
Under uniform convexity of $W$ and the linearity of
$\bs q\mapsto-\bs q\cdot\bs\ell(S)$, the reduced energy $E(\cdot,S)$ is
strictly (indeed uniformly) convex, with
$D^2_{\bs q}E(\bs q,S)=D^2W(\bs q)\succeq\alpha\Id$ for every $S\ge0$.
Strict convexity makes the global stability $\bs q_0\in\mathcal S(0)$
equivalent to the pointwise first-order inclusion
$-\partial_{\bs q}E(\bs q_0,S(0))\in\partial\Psi(\bs 0)$, i.e.\ the
initial state is a stationary point of the incremental potential. By the
strengthened smoothness $\bs\ell\in C^3$, $S\in C^3([0,T])$, the map $E$
is jointly $C^3$ on $[0,T]\times\Q$. These are exactly the hypotheses of
\cite[Thm.~3.4.7]{MielkeRoubicek}, whose Section~3.4.4 establishes that
the energetic solution emanating from $\bs q_0$ is unique.
\end{proof}

\subsection{Balanced-viscosity solutions}

\begin{proof}[Proof of \cref{thm:BV_existence}]
\emph{Statement.} \emph{Under \ref{A:W}--\ref{A:coerc}, for every
$\bs q_0\in\mathcal S(0)$ there exists at least one BV solution in the
sense of \cref{def:BV}.}

\smallskip
\emph{Step 1: the regularised problem.} For $\varepsilon>0$ write
$\Phi_\varepsilon(\bs v):=\Psi(\bs v)+\tfrac{\varepsilon}{2}\|\bs v\|^2$;
then \eqref{eq:regularised} is $\bs 0\in\partial\Phi_\varepsilon(\dot{\bs q}_\varepsilon)+\partial_{\bs q}E(\bs q_\varepsilon,S)$.
As $\Phi_\varepsilon$ is strictly convex, coercive and continuous, and
$\partial_{\bs q}E(\cdot,S)$ is a $C^1$ perturbation, the theory of
doubly-nonlinear evolution inclusions governed by a maximal monotone
dissipation \cite{Brezis1973} yields, for each $\varepsilon>0$, a unique
solution $\bs q_\varepsilon\in W^{1,2}(0,T;\Q)$ with
$\bs q_\varepsilon(0)=\bs q_0$, satisfying the (strict) energy identity
\begin{equation}\label{eq:reg_energy}
  E(\bs q_\varepsilon(t),S(t))+\int_0^t\Bigl[\Psi(\dot{\bs q}_\varepsilon(s))+\varepsilon\|\dot{\bs q}_\varepsilon(s)\|^2\Bigr]\dd s
  =E(\bs q_0,S(0))+\int_0^t\partial_S E(\bs q_\varepsilon,S(s))\dot S(s)\dd s .
\end{equation}

\emph{Step 2: uniform energy and total-variation bounds.} Dropping the
non-negative viscous term and the non-negative dissipation
in~\eqref{eq:reg_energy}, and abbreviating
$\phi_\varepsilon(t):=E(\bs q_\varepsilon(t),S(t))+c_4$ (with $c_4$ from
\cref{prop:power}\ref{E:gronwall}, so $\phi_\varepsilon\ge1$),
\[
  \phi_\varepsilon(t)\le\phi_\varepsilon(0)+\int_0^t|\partial_S E(\bs q_\varepsilon,S(s))|\,|\dot S(s)|\dd s
  \le\phi_\varepsilon(0)+\int_0^t c_3\,\phi_\varepsilon(s)\,|\dot S(s)|\dd s,
\]
using \cref{prop:power}\ref{E:gronwall}. Since $|\dot S|\in L^1$,
Gronwall's inequality gives
$\phi_\varepsilon(t)\le\phi_\varepsilon(0)\exp\!\bigl(c_3\|\dot S\|_{L^1}\bigr)$,
whence $\sup_t E(\bs q_\varepsilon(t),S(t))\le M$ with $M$ independent of
$\varepsilon$; by coercivity \ref{A:W}, $\sup_t\|\bs q_\varepsilon(t)\|\le M'$,
also independent of $\varepsilon$. Now, using \ref{A:coerc}
($\Psi(\bs v)\ge c_\Psi\|\bs v\|$), \cref{lem:diss_ac} and the energy
identity,
\begin{align}
  c_\Psi\,\Var(\bs q_\varepsilon;[0,T])
  &=c_\Psi\int_0^T\|\dot{\bs q}_\varepsilon(s)\|\dd s
  \le\int_0^T\Psi(\dot{\bs q}_\varepsilon(s))\dd s
  =\Diss_\Psi(\bs q_\varepsilon;[0,T]) \notag\\
  &\le E(\bs q_0,S(0))-E(\bs q_\varepsilon(T),S(T))
  +\int_0^T|\partial_S E(\bs q_\varepsilon,S(s))|\,|\dot S(s)|\dd s \notag\\
  &\le E(\bs q_0,S(0))+C_{\min}+C_{\ell'}M'\|\dot S\|_{L^1}
  =:c_\Psi\,C, \label{eq:var_bound_bv}
\end{align}
where $E(\bs q_\varepsilon(T),S(T))\ge-C_{\min}$ by coercivity and
$|\partial_S E|\le C_{\ell'}\|\bs q_\varepsilon\|\le C_{\ell'}M'$. Thus
$\Var(\bs q_\varepsilon;[0,T])\le C$, uniformly in $\varepsilon$.

\emph{Step 3: vanishing-viscosity limit.} By Helly's selection theorem
\cite[Thm.~3.3]{AmbrosioFuscoPallara}, the uniform bounds
$\|\bs q_\varepsilon\|_{L^\infty}\le M'$ and $\Var(\bs q_\varepsilon)\le C$
give a subsequence $\varepsilon_j\to0^+$ and a limit
$\bs q\in\mathrm{BV}([0,T];\Q)$ with $\bs q_{\varepsilon_j}(t)\to\bs q(t)$
for every $t\in[0,T]$. Local stability (BV-S) passes to the limit by
lower semicontinuity of $E(\cdot,S)$ (continuity of $W$ and $\bs\ell$)
and continuity of $\Dr$ (\cref{prop:Dr_linear}). For the work term in
(BV-E), the integrands
$\partial_S E(\bs q_{\varepsilon_j}(s),S(s))\dot S(s)=-\bs q_{\varepsilon_j}(s)\cdot\bs\ell'(S(s))\dot S(s)$
converge pointwise a.e.\ to $\partial_S E(\bs q(s),S(s))\dot S(s)$ and
are dominated, uniformly in $j$, by the integrable function
$s\mapsto C_{\ell'}M'|\dot S(s)|\in L^1(0,T)$; Lebesgue's dominated
convergence theorem gives the convergence of the work integral. The
identification of the viscous correction $\mathcal V$ as the limit of
$\int\varepsilon\|\dot{\bs q}_\varepsilon(s)\|^2 \dd s$ concentrated at the jumps
of $\bs q$ follows from the reparametrised (arc-length) passage to the
limit of \cite[Thm.~3.3.2]{MielkeRoubicek}. Hence $\bs q$ is a BV
solution in the sense of \cref{def:BV}.
\end{proof}

\begin{proof}[Proof of \cref{thm:BV_uniqueness}]
\emph{Statement.} \emph{Under \ref{A:W}--\ref{A:coerc}, if $n=1$ and
$q\mapsto E(q,S)$ has finitely many local minima for each $S$, the BV
solution is unique.}

\smallskip
\emph{Uniqueness of the post-jump state.} At a jump time $t^\ast$, the
pre-jump state $q^-(t^\ast)$ is determined by left-continuity of the
preceding arc. In dimension one, any monotone path from $q^-$ to a
candidate $q^+$ realises the dissipation distance $\Dr(q^-,q^+)=\rho|q^+-q^-|$
(\cref{prop:Dr_linear}), and the balanced-viscosity jump condition
extracted from (BV-E) --- namely that the transition follows the
vanishing-viscosity heteroclinic, i.e.\ the path of steepest descent of
$q\mapsto E(q,S(t^\ast))+\rho|q-q^-|$ --- selects $q^+$ as the first
local minimiser of $E(\cdot,S(t^\ast))$ encountered in the direction of
motion beyond the point where $q^-$ loses local stability. Because
$E(\cdot,S(t^\ast))$ has finitely many local minima, this ``first''
minimiser is unambiguous; hence $q^+(t^\ast)$ is unique.

\emph{Uniqueness of the trajectory.} On any interval free of jumps, the
solution is locally stable and satisfies the flow rule
$f\in\partial\Psi(\dot q)$, i.e.\ the scalar Kuhn--Tucker system
$|f|\le\rho$ with $\dot q\ne0$ only when $|f|=\rho$; this determines
$q(t)$ uniquely as the monotone solution of the algebraic yield
equation $g(q)=\ell(S)\mp\rho$ on the current branch, with the sign
fixed by the direction of loading. Since both the smooth arcs and the
jump targets are uniquely determined and the jump times are the
(isolated) instants at which local stability is lost, the whole
trajectory is unique.
\end{proof}

% =====================================================================
\section{Variational numerical method}
\label{app:numerical}

\begin{proposition}[Return map]\label{prop:return_map}
Under \ref{A:Psi}, the incremental problem~\eqref{eq:Ikstar} is
equivalent to the inclusion~\eqref{eq:discrete_inclusion}, which
decomposes into two mutually exclusive cases:
\textnormal{(i)} \emph{elastic lock}, $\bs q_{k+1}=\bs q_k$, iff
$-\partial_{\bs q}E(\bs q_k,S_{k+1})\in\E$; and
\textnormal{(ii)} \emph{dissipative flow}, $\bs q_{k+1}\ne\bs q_k$ with
$-\partial_{\bs q}E(\bs q_{k+1},S_{k+1})\in\partial\Psi(\bs q_{k+1}-\bs q_k)\setminus\E$.
In the scalar case $\Psi(\dot q)=\rho|\dot q|$ this is precisely the
return map~\eqref{eq:scalar_return_map}.
\end{proposition}

\begin{proof}
The functional $\bs q\mapsto E(\bs q,S_{k+1})+\Dr(\bs q_k,\bs q)$ is the
sum of the differentiable $E(\cdot,S_{k+1})$ and the convex
$\Dr(\bs q_k,\cdot)$ (\cref{prop:quasi_distance}); hence $\bs q_{k+1}$
minimises it iff
\begin{equation}\label{eq:opt_incremental}
  \bs 0\in\partial_{\bs q}E(\bs q_{k+1},S_{k+1})+\partial_{\bs q}\Dr(\bs q_k,\bs q_{k+1}).
\end{equation}
By \cref{prop:Dr_linear}, $\Dr(\bs q_k,\bs q)=\Psi(\bs q-\bs q_k)$, so
$\partial_{\bs q}\Dr(\bs q_k,\bs q)=\partial\Psi(\bs q-\bs q_k)$ and
\eqref{eq:opt_incremental} is exactly~\eqref{eq:discrete_inclusion}. If
$\bs q_{k+1}=\bs q_k$, then $\bs q_{k+1}-\bs q_k=\bs 0$ and the inclusion
reads $-\partial_{\bs q}E(\bs q_k,S_{k+1})\in\partial\Psi(\bs 0)=\E$,
giving case~(i). If $\bs q_{k+1}\ne\bs q_k$, then
$-\partial_{\bs q}E(\bs q_{k+1},S_{k+1})\in\partial\Psi(\bs q_{k+1}-\bs q_k)$,
which by \cref{lem:saturation}\ref{sat:ii} lies in $\E$ only on its
boundary; case~(ii) records that the force is on $\partial\E$ and drives
a non-zero increment. In the scalar case, $\partial\Psi(v)=\{\rho\}$ for
$v>0$, $[-\rho,\rho]$ for $v=0$, $\{-\rho\}$ for $v<0$; writing the
trial force $f^{\mathrm{tr}}=\ell(S_{k+1})-g(q_k)$, case~(i) is
$|f^{\mathrm{tr}}|\le\rho$, while case~(ii) forces
$\ell(S_{k+1})-g(q_{k+1})=\pm\rho$ with the sign of $q_{k+1}-q_k$, i.e.\
$g(q_{k+1})=\ell(S_{k+1})-\sgn(f^{\mathrm{tr}})\rho$, which
is~\eqref{eq:scalar_return_map}.
\end{proof}

\begin{proposition}[Discrete stability and energy inequality]
\label{prop:discrete_energy}
Let $\{\bs q_k\}_{k=0}^N$ be generated by~\eqref{eq:Ikstar}. Then, for
each $k$,
\begin{align}
  &E(\bs q_{k+1},S_{k+1})\le E(\tilde{\bs q},S_{k+1})+\Dr(\bs q_k,\tilde{\bs q})
  \qquad\forall\,\tilde{\bs q}\in\Q, \label{eq:disc_stab}\\
  &E(\bs q_{k+1},S_{k+1})+\Dr(\bs q_k,\bs q_{k+1})\le E(\bs q_k,S_{k+1}). \label{eq:disc_energy}
\end{align}
\end{proposition}

\begin{proof}
By minimality of $\bs q_{k+1}$ in~\eqref{eq:Ikstar}, for every
$\tilde{\bs q}\in\Q$,
\begin{equation}\label{eq:min_incremental}
  E(\bs q_{k+1},S_{k+1})+\Dr(\bs q_k,\bs q_{k+1})
  \le E(\tilde{\bs q},S_{k+1})+\Dr(\bs q_k,\tilde{\bs q}).
\end{equation}
Since $\Dr(\bs q_k,\bs q_{k+1})\ge0$ (D1), dropping it from the
left-hand side of~\eqref{eq:min_incremental} gives~\eqref{eq:disc_stab}.
Taking $\tilde{\bs q}=\bs q_k$ in~\eqref{eq:min_incremental} and using
$\Dr(\bs q_k,\bs q_k)=0$ (D1) gives~\eqref{eq:disc_energy}.
\end{proof}

\begin{proof}[Proof of \cref{thm:convergence}]
\emph{Statement.} \emph{Under \ref{A:W}--\ref{A:Psi}, with
$\bs q_0\in\mathcal S(0)$ and $\bs q^h$ the piecewise-constant
interpolant of the incremental sequence on a partition of maximal step
$h$, a subsequence $\bs q^{h_j}(t)\to\bs q(t)$ for every $t$, with
$\bs q$ an energetic solution; under uniqueness the whole family
converges.}

\smallskip
Write $\alpha_k:=\int_{t_k}^{t_{k+1}}|\dot S(s)|\dd s$, so that
$\sum_k\alpha_k=\|\dot S\|_{L^1(0,T)}$.

\emph{Step 1: uniform energy and variation bounds.} Summing the
discrete energy inequality~\eqref{eq:disc_energy} over $k=0,\dots,K-1$
telescopes the stored-energy terms to
\begin{equation}\label{eq:summed_energy}
  E(\bs q_K,S_K)+\sum_{k=0}^{K-1}\Dr(\bs q_k,\bs q_{k+1})
  \le E(\bs q_0,S_0)+\sum_{k=0}^{K-1}\bigl[E(\bs q_k,S_{k+1})-E(\bs q_k,S_k)\bigr].
\end{equation}
By \cref{prop:power}\ref{E:AC} and Cauchy--Schwarz, each work increment
is controlled:
\begin{equation}\label{eq:work_incr}
  E(\bs q_k,S_{k+1})-E(\bs q_k,S_k)
  =\int_{t_k}^{t_{k+1}}\!(-\bs q_k\cdot\bs\ell'(S(s)))\dot S(s)\dd s
  \le C_{\ell'}\|\bs q_k\|\,\alpha_k .
\end{equation}
The coercivity \ref{A:W} together with \eqref{eq:young} (as in
\cref{prop:power}\ref{E:gronwall}) gives, with $C_p:=(2/c_1)^{1/p}$ and
$\tilde c_2:=c_2+K$,
\begin{equation}\label{eq:qk_linear}
  \|\bs q_k\|\le C_p\bigl(E(\bs q_k,S_k)+\tilde c_2+1\bigr).
\end{equation}
Substituting~\eqref{eq:work_incr} and~\eqref{eq:qk_linear} into
\eqref{eq:summed_energy} and dropping the non-negative dissipation sum,
and setting $e_k:=E(\bs q_k,S_k)+\tilde c_2+1$,
\begin{equation}\label{eq:disc_gronwall}
  e_K\le e_0+C_{\ell'}C_p\sum_{k=0}^{K-1}e_k\,\alpha_k .
\end{equation}
The discrete Gronwall lemma\footnote{If $e_K\le e_0+C\sum_{k<K}e_k\alpha_k$
with $e_k,\alpha_k\ge0$, then $e_K\le e_0\exp\!\bigl(C\sum_{k<K}\alpha_k\bigr)$;
proved by induction, comparing with the solution of the corresponding
equality.}
applied to~\eqref{eq:disc_gronwall} yields
$e_K\le e_0\exp\!\bigl(C_{\ell'}C_p\|\dot S\|_{L^1}\bigr)=:\widetilde M$,
so $\sup_{k,h}E(\bs q_k,S_k)\le\widetilde M-\tilde c_2-1=:M$, a bound
independent of $h$. Coercivity then gives
$c_1\|\bs q_k\|^p-C_\ell\|\bs q_k\|\le M+c_2$, whence
$\sup_{k,h}\|\bs q_k\|\le M'$ independent of $h$. Finally,
from~\eqref{eq:summed_energy} with $K=N$, using
$E(\bs q_N,S_N)\ge-c_2-C_\ell M'=:-C_{\min}$ and \eqref{eq:work_incr},
\begin{equation}\label{eq:var_bound}
  \Var(\bs q^h;[0,T])=\sum_{k=0}^{N-1}\Dr(\bs q_k,\bs q_{k+1})
  \le E(\bs q_0,S_0)+C_{\min}+C_{\ell'}M'\|\dot S\|_{L^1}=:C_{\mathrm{var}},
\end{equation}
with $C_{\mathrm{var}}$ independent of $h$.

\emph{Step 2: compactness.} By~\eqref{eq:var_bound} and
$\|\bs q^h\|_{L^\infty}\le M'$, Helly's selection theorem
\cite[Thm.~3.3]{AmbrosioFuscoPallara} yields a subsequence $h_j\to0$ and
$\bs q\in\mathrm{BV}([0,T];\Q)$ with $\bs q^{h_j}(t)\to\bs q(t)$ for
every $t\in[0,T]$; $\bs q(t)\in\Q$ since $\Q$ is closed.

\emph{Step 3: verification of \textnormal{(S)}.} Fix $t$ and
$\tilde{\bs q}\in\Q$. Let $k_j$ be the index with $t\in[t_{k_j},t_{k_j+1})$;
then $t_{k_j+1}\to t$ and, by continuity of $S$ ($S\in W^{1,1}\subset C^0$),
$S_{k_j+1}\to S(t)$. The discrete stability~\eqref{eq:disc_stab} at step
$k_j$ reads
$E(\bs q^{h_j}(t),S_{k_j+1})\le E(\tilde{\bs q},S_{k_j+1})+\Dr(\bs q^{h_j}(t),\tilde{\bs q})$.
Passing to the limit $j\to\infty$: the left side tends to
$E(\bs q(t),S(t))$ and the first right-hand term to
$E(\tilde{\bs q},S(t))$ by continuity of $E$ and $\bs q^{h_j}(t)\to\bs q(t)$,
$S_{k_j+1}\to S(t)$; the second term tends to $\Dr(\bs q(t),\tilde{\bs q})$
by continuity of $\Dr$ (\cref{prop:Dr_linear}). Hence
$E(\bs q(t),S(t))\le E(\tilde{\bs q},S(t))+\Dr(\bs q(t),\tilde{\bs q})$
for all $\tilde{\bs q}$, i.e.\ (S).

\emph{Step 4: verification of \textnormal{(E)}.} \textit{Upper bound.}
With $K_j$ such that $t_{K_j}\le t<t_{K_j+1}$, summing
\eqref{eq:disc_energy} over $k<K_j$ and writing the work exactly, as
in~\eqref{eq:work_incr} but with equality, gives
\begin{equation}\label{eq:summed_exact}
  E(\bs q_{K_j},S_{K_j})+\Diss_\Psi(\bs q^{h_j};[0,t_{K_j}])
  \le E(\bs q_0,S_0)+\int_0^{t_{K_j}}\!\partial_S E(\bs q^{h_j}(s),S(s))\dot S(s)\dd s,
\end{equation}
where we used $\bs q^{h_j}(s)=\bs q_k$ on $[t_k,t_{k+1})$. We pass to the
limit in each term. On the left, $E(\bs q_{K_j},S_{K_j})\to E(\bs q(t),S(t))$
by continuity, while lower semicontinuity of the $\Psi$-variation under
pointwise convergence gives
$\Diss_\Psi(\bs q;[0,t])\le\liminf_j\Diss_\Psi(\bs q^{h_j};[0,t])$: for
any fixed partition $0=s_0<\dots<s_M=t$,
$\sum_i\Dr(\bs q^{h_j}(s_i),\bs q^{h_j}(s_{i+1}))\to\sum_i\Dr(\bs q(s_i),\bs q(s_{i+1}))$
by continuity of $\Dr$, and each partial sum is bounded by
$\Diss_\Psi(\bs q^{h_j};[0,t])$; taking $\liminf_j$ and then the
supremum over partitions gives the claim. On the right, the integrands
$\partial_S E(\bs q^{h_j}(s),S(s))\dot S(s)=-\bs q^{h_j}(s)\cdot\bs\ell'(S(s))\dot S(s)$
converge pointwise a.e.\ to $\partial_S E(\bs q(s),S(s))\dot S(s)$ (as
$\bs q^{h_j}(s)\to\bs q(s)$ and $\bs\ell',\dot S$ are fixed) and are
dominated, uniformly in $j$, by the integrable function
$s\mapsto C_{\ell'}M'|\dot S(s)|\in L^1(0,T)$ (using
$\|\bs q^{h_j}(s)\|\le M'$ and $\|\bs\ell'\|\le C_{\ell'}$); Lebesgue's
dominated convergence theorem therefore gives
$\int_0^{t_{K_j}}\partial_S E(\bs q^{h_j},S)\dot S\to\int_0^t\partial_S E(\bs q,S)\dot S$.
Therefore
\begin{equation}\label{eq:E_upper}
  E(\bs q(t),S(t))+\Diss_\Psi(\bs q;[0,t])
  \le E(\bs q_0,S(0))+\int_0^t\partial_S E(\bs q,S(s))\dot S(s)\dd s .
\end{equation}
\textit{Lower bound.} Since $\bs q$ satisfies (S) (Step~3), the standard
lower energy estimate for stable BV curves holds: testing (S) along any
partition $0=s_0<\dots<s_M=t$, summing the resulting inequalities
$E(\bs q(s_{i}),S(s_i))\le E(\bs q(s_{i+1}),S(s_i))+\Dr(\bs q(s_{i}),\bs q(s_{i+1}))$
and passing to the finest partition recovers the reverse inequality
of~\eqref{eq:E_upper} (see \cite[Sec.~2.1]{MielkeRoubicek}).
Combining the two inequalities yields the energy balance~(E). Hence
$\bs q$ is an energetic solution. If the solution is unique
(\cref{thm:uniqueness} or the scalar case of \cref{thm:BV_uniqueness}),
every subsequence has the same limit $\bs q$, so the whole family
$\bs q^h$ converges.
\end{proof}

\begin{proposition}[Local consistency]\label{prop:consistency}
Assume \ref{A:ell} and $S\in W^{1,\infty}(0,T)$, and let $\bs q$ be the
exact energetic solution and $\tau_k$ the truncation
error~\eqref{eq:truncation}.
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=2.4em,itemsep=2pt]
  \item\label{cons:weak} If $\bs q\in\mathrm{AC}(0,T;\Q)$, then
  \begin{equation}\label{eq:cons_weak}
    |\tau_k|\le C_{\ell'}\|\dot S\|_{L^\infty}\Bigl(\int_{t_k}^{t_{k+1}}\|\dot{\bs q}(\tau)\|\dd\tau\Bigr)h_k=O(h_k).
  \end{equation}
  \item\label{cons:strong} If $\bs q\in W^{1,\infty}(0,T;\Q)$
  \textnormal{(}equivalently, $\bs q$ Lipschitz\textnormal{)}, then
  \begin{equation}\label{eq:cons_strong}
    |\tau_k|\le\tfrac12 C_{\ell'}\|\dot{\bs q}\|_{L^\infty}\|\dot S\|_{L^\infty}\,h_k^2=O(h_k^2).
  \end{equation}
\end{enumerate}
\end{proposition}

\begin{proof}
By the fundamental theorem of calculus (valid since
$S\in W^{1,\infty}$), the frozen-state discrete work is
$E(\bs q(t_k),S_{k+1})-E(\bs q(t_k),S_k)=\int_{t_k}^{t_{k+1}}\partial_S E(\bs q(t_k),S(s))\dot S(s)\dd s$.
Subtracting this from the continuous work in~\eqref{eq:truncation} and
using $\partial_S E(\bs q,S)=-\bs q\cdot\bs\ell'(S)$,
\begin{align}
  \tau_k
  &=\int_{t_k}^{t_{k+1}}\bigl[\partial_S E(\bs q(s),S(s))-\partial_S E(\bs q(t_k),S(s))\bigr]\dot S(s)\dd s \notag\\
  &=\int_{t_k}^{t_{k+1}}\bigl(\bs q(s)-\bs q(t_k)\bigr)\cdot\bigl(-\bs\ell'(S(s))\bigr)\dot S(s)\dd s . \label{eq:tau_repr}
\end{align}
By Cauchy--Schwarz, \ref{A:ell} ($\|\bs\ell'\|\le C_{\ell'}$) and
$|\dot S|\le\|\dot S\|_{L^\infty}$,
\begin{equation}\label{eq:tau_bound_common}
  |\tau_k|\le C_{\ell'}\|\dot S\|_{L^\infty}\int_{t_k}^{t_{k+1}}\|\bs q(s)-\bs q(t_k)\|\dd s .
\end{equation}
\emph{\ref{cons:weak}.} If $\bs q\in\mathrm{AC}$, then for
$s\in[t_k,t_{k+1}]$, $\|\bs q(s)-\bs q(t_k)\|\le\int_{t_k}^{s}\|\dot{\bs q}\|\dd\tau\le\int_{t_k}^{t_{k+1}}\|\dot{\bs q}\|\dd\tau$;
inserting into~\eqref{eq:tau_bound_common} and integrating the constant
bound over $s\in[t_k,t_{k+1}]$ (length $h_k$) gives~\eqref{eq:cons_weak}.
\emph{\ref{cons:strong}.} If $\bs q\in W^{1,\infty}$, i.e.\ Lipschitz,
then $\|\bs q(s)-\bs q(t_k)\|\le\|\dot{\bs q}\|_{L^\infty}(s-t_k)$, so
\eqref{eq:tau_bound_common} gives
$|\tau_k|\le C_{\ell'}\|\dot S\|_{L^\infty}\|\dot{\bs q}\|_{L^\infty}\int_{t_k}^{t_{k+1}}(s-t_k)\dd s
=\tfrac12 C_{\ell'}\|\dot S\|_{L^\infty}\|\dot{\bs q}\|_{L^\infty}h_k^2$,
which is~\eqref{eq:cons_strong}.
\end{proof}

\begin{proof}[Proof of \cref{thm:global_consistency}]
\emph{Statement.} \emph{Under \ref{A:ell}, $S\in W^{1,\infty}$ and
either $\bs q\in\mathrm{AC}$ or $\bs q\in W^{1,\infty}$, the accumulated
truncation error satisfies $\mathcal E_\tau=\sum_k|\tau_k|\le Ch$.}

\smallskip
Sum the local bounds of \cref{prop:consistency} over $k=0,\dots,N-1$ and
bound each step by $h=\max_k h_k$. In the weak case,
by~\eqref{eq:cons_weak},
\[
  \mathcal E_\tau=\sum_{k}|\tau_k|
  \le C_{\ell'}\|\dot S\|_{L^\infty}\sum_k h_k\int_{t_k}^{t_{k+1}}\|\dot{\bs q}\|\dd\tau
  \le C_{\ell'}\|\dot S\|_{L^\infty}\,h\sum_k\int_{t_k}^{t_{k+1}}\|\dot{\bs q}\|\dd\tau
  = C_{\ell'}\|\dot S\|_{L^\infty}\,h\!\int_0^T\!\|\dot{\bs q}\|\dd\tau,
\]
the last equality because the subinterval integrals partition $[0,T]$.
In the strong case, by~\eqref{eq:cons_strong} and $\sum_k h_k^2\le h\sum_k h_k=Th$,
\[
  \mathcal E_\tau\le\tfrac12 C_{\ell'}\|\dot{\bs q}\|_{L^\infty}\|\dot S\|_{L^\infty}\sum_k h_k^2
  \le\tfrac12 C_{\ell'}\|\dot{\bs q}\|_{L^\infty}\|\dot S\|_{L^\infty}\,T\,h .
\]
In both cases $\mathcal E_\tau\le C\,h=O(h)$, with $C$ as stated in
\cref{thm:global_consistency}.
\end{proof}

\begin{proposition}[State error under nodal exactness]\label{prop:interp}
Let $\bs q\colon[0,T]\to\Q$ have finitely many jump points
$0<t^\ast_1<\dots<t^\ast_m<T$, be Lipschitz with constant $L$ on each
open subinterval between consecutive jumps, and set
$\Delta_j:=\|\bs q(t^{\ast+}_j)-\bs q(t^{\ast-}_j)\|$ and
$J:=\sum_{j=1}^m\Delta_j$. On a partition of maximal step $h$, let
$\bs q^h$ be the right-continuous piecewise-constant interpolant,
$\bs q^h(t)=\bs q^h(t_i)$ for $t\in[t_i,t_{i+1})$, and assume the scheme
is \emph{node-exact}: $\bs q^h(t_i)=\bs q(t_i)$ at every node $t_i$. Then
\begin{enumerate}[label=\textnormal{(\roman*)},leftmargin=2.4em,itemsep=2pt]
  \item\label{interp:sup} for every interval $I\subset[0,T]$ at distance
  greater than $h$ from $\{t^\ast_1,\dots,t^\ast_m\}$,
  $\|\bs q^h-\bs q\|_{L^\infty(I)}\le L\,h$;
  \item\label{interp:L1} $\|\bs q^h-\bs q\|_{L^1(0,T;\Q)}\le(\tfrac12 LT+J)\,h+mL\,h^2$.
\end{enumerate}
In particular $\|\bs q^h-\bs q\|_{L^1}=O(h)$, and the sup-norm error away
from the jumps is $O(h)$.
\end{proposition}

\begin{proof}
Fix a step $[t_i,t_{i+1})$, $h_i:=t_{i+1}-t_i\le h$. Node-exactness gives
$\bs q^h\equiv\bs q(t_i)$ on the step, so
$\|\bs q^h(t)-\bs q(t)\|=\|\bs q(t_i)-\bs q(t)\|$ there.

\emph{Jump-free steps.} If the step contains no jump, $\bs q$ is
$L$-Lipschitz on it, so $\|\bs q(t_i)-\bs q(t)\|\le L(t-t_i)\le L h_i\le Lh$.
This proves~\ref{interp:sup}, because an interval at distance $>h$ from
the jumps meets only jump-free steps (each step has length $\le h$).
Moreover $\int_{t_i}^{t_{i+1}}\|\bs q^h-\bs q\|\dd t\le\int_{t_i}^{t_{i+1}}L(t-t_i)\dd t=\tfrac12 L h_i^2$.

\emph{Steps containing a jump $t^\ast_j$.} The total variation of
$\bs q$ from $t_i$ to $t$ is its Lipschitz part plus the single crossed
jump, so $\|\bs q(t_i)-\bs q(t)\|\le L(t-t_i)+\Delta_j\le L h_i+\Delta_j$,
whence $\int_{t_i}^{t_{i+1}}\|\bs q^h-\bs q\|\dd t\le(L h_i+\Delta_j)h_i\le L h_i^2+\Delta_j\,h$.

\emph{Summation.} Adding over all steps, using $\sum_i h_i=T$,
$h_i\le h$, and that at most $m$ steps contain a jump,
\[
  \|\bs q^h-\bs q\|_{L^1}
  \le\tfrac12 L\sum_i h_i^2+\sum_{j=1}^m\bigl(L h_{i(j)}^2+\Delta_j\,h\bigr)
  \le\tfrac12 L T h+mL\,h^2+J\,h,
\]
which is~\ref{interp:L1}; both bounds are $O(h)$.
\end{proof}

\begin{remark}\label{rem:interp_scope}
\Cref{prop:interp} presumes the node-exactness $\bs q^h(t_i)=\bs q(t_i)$,
which is \emph{not} a generic property of the scheme; it holds in the
scalar benchmarks of \cref{sec:experiments} because, on any interval of
monotone loading, the return map solves the same algebraic yield
condition that the exact solution satisfies at the node, and reproduces
the lock values exactly. It fails whenever the loading carries structure
on a time scale comparable to, or finer than, the step $h$---so that the
exact state changes regime more than once within a step---the simplest
instance being a loading extremum strictly interior to a step, and a more
striking one a sub-step oscillation, for which the state error remains
$O(1)$ until $h$ resolves that scale (\cref{sec:conv_osc}). It fails
likewise, in the vectorial case, where the transition path at a jump is
genuinely path-dependent and the discrete and continuous post-jump states
may differ. A general error rate without node-exactness is not addressed
here.
\end{remark}

% =====================================================================
\section{Details for the bistable benchmark}
\label{app:experiments}

We prove the energetic jump threshold~\eqref{eq:l_jump_energetic} of
\cref{sec:benchmark2} and record the local-stability threshold quoted in
\cref{rem:energetic_vs_bv}. Throughout, $k,a,\rho>0$ and $\ell\ge0$ is
the (increasing) instantaneous load; we use the incremental
objective~\eqref{eq:Ikstar} with previous state $q_k=-a$,
\begin{equation}\label{eq:I_bistable}
  I(q):=E(q,\ell)+\rho\,|q-(-a)|=W(q)-q\ell+\rho\,|q+a|,
\end{equation}
with $W$ the piecewise-quadratic well~\eqref{eq:benchmark2}. Condition
(S) at $q=-a$ holds iff $-a$ is a global minimiser of $I$ over $\R$.

\begin{proposition}[Energetic jump threshold]\label{prop:jump_energetic}
For~\eqref{eq:benchmark2} with initial datum $q(0)=-a$ and increasing
loading, and provided $\rho<4ka$, the state $q=-a$ is globally stable
exactly on $\{\ell:\ell\le\rho\}$ and loses global stability for
$\ell>\rho$, the comparison being realised by the bottom of the other
well, $q=+a$. Equivalently, the energetic solution jumps at
$\ell_{\mathrm{jump}}=\rho$, from $q^-=-a$ to $q^+=+a$.
\end{proposition}

\begin{proof}
We minimise $I$ in~\eqref{eq:I_bistable} separately on the two branches
$q\le0$ and $q>0$, each of which is strictly convex, then compare their
minima; the switch occurs where the two are equal, a \emph{Maxwell}
(equal-area) construction in the sense of phase-transition
theory~\cite{mielke2002}: the globally stable branch is the one of lower
incremental energy, and the transition occurs where the two branches
are energetically balanced.

\emph{Left branch ($q\le0$).} Here $W(q)=\tfrac12k(q+a)^2$ and
$|q+a|=q+a$ for $q \ge -a$ (note that $I(q)$ is strictly decreasing for $q < -a$, so the minimiser in the left branch must satisfy $q \ge -a$), so
$I(q)=\tfrac12k(q+a)^2-q\ell+\rho(q+a)$ with
$I'(q)=k(q+a)-\ell+\rho$. At $q=-a$, $I'(-a)=\rho-\ell$, which is
$\ge0$ precisely when $\ell\le\rho$; since $I$ is convex on this branch,
$q=-a$ is its minimiser over $\{q\le0\}$ for all $\ell\le\rho$, with
minimum value
\begin{equation}\label{eq:I_left}
  I(-a)=W(-a)+a\ell=a\ell
\end{equation}
(the quadratic term vanishes at the well bottom $q=-a$).

\emph{Right branch ($q>0$).} Here $W(q)=\tfrac12k(q-a)^2$ and
$|q+a|=q+a$, so $I(q)=\tfrac12k(q-a)^2-q\ell+\rho(q+a)$ with unconstrained
minimiser $q_R:=a+(\ell-\rho)/k$ (positive once $\ell>\rho-ka$). Using
$q_R-a=(\ell-\rho)/k$,
\[
  W(q_R)=\frac{(\ell-\rho)^2}{2k},\qquad
  E(q_R,\ell)=W(q_R)-q_R\ell
  =\frac{(\ell-\rho)^2}{2k}-a\ell-\frac{\ell(\ell-\rho)}{k},
\]
and, since $q_R+a=2a+(\ell-\rho)/k$,
\begin{equation}\label{eq:I_right}
  I(q_R)=E(q_R,\ell)+\rho(q_R+a)
  =\frac{(\ell-\rho)^2}{2k}-a\ell-\frac{\ell(\ell-\rho)}{k}+2a\rho+\frac{\rho(\ell-\rho)}{k}.
\end{equation}
The three $\tfrac1k$-terms in~\eqref{eq:I_right} combine as
$\tfrac1k\bigl[\tfrac12(\ell-\rho)^2-\ell(\ell-\rho)+\rho(\ell-\rho)\bigr]
=\tfrac1k(\ell-\rho)\bigl[\tfrac12(\ell-\rho)-(\ell-\rho)\bigr]
=-\tfrac{(\ell-\rho)^2}{2k}$, so
\begin{equation}\label{eq:I_right_simplified}
  I(q_R)=-\frac{(\ell-\rho)^2}{2k}-a\ell+2a\rho .
\end{equation}

\emph{Comparison.} Subtracting~\eqref{eq:I_right_simplified}
from~\eqref{eq:I_left},
\begin{equation}\label{eq:maxwell}
  I(-a)-I(q_R)
  =a\ell-\Bigl(-\frac{(\ell-\rho)^2}{2k}-a\ell+2a\rho\Bigr)
  =2a(\ell-\rho)+\frac{(\ell-\rho)^2}{2k}
  =(\ell-\rho)\Bigl[2a+\frac{\ell-\rho}{2k}\Bigr].
\end{equation}
The bracket $2a+(\ell-\rho)/(2k)$ is increasing in $\ell$; over
$\ell\in[0,\rho]$ its minimum is at $\ell=0$, equal to
$2a-\rho/(2k)$, which is positive precisely when $\rho<4ka$. Under this
hypothesis the bracket is positive throughout $[0,\rho]$, so the sign of
\eqref{eq:maxwell} is that of $(\ell-\rho)$: it is $\le0$ for
$\ell\le\rho$ (with equality only at $\ell=\rho$) and $>0$ for $\ell$
just above $\rho$ (the bracket stays positive there). Consequently
$I(-a)\le I(q_R)$ for all $\ell\in[0,\rho]$, and since $q_R$ is $I$'s
minimiser on the right branch while $-a$ is $I$'s minimiser on the left
branch, $q=-a$ is the global minimiser of $I$ over $\R$ exactly for
$\ell\in[0,\rho]$; it ceases to be so immediately beyond $\ell=\rho$. At
$\ell=\rho$ one has $q_R=a$, so the losing comparison is with $q=+a$,
proving the claim.
\end{proof}

\begin{remark}[Local threshold]\label{rem:local_threshold}
The distinction invoked in \cref{rem:energetic_vs_bv} concerns not the
pre-jump state but \emph{how long the left branch remains locally
admissible}. Tracking the left-branch equilibrium
$q_L(\ell)=-a+(\ell-\rho)/k$ (the stationary point of $I$ on
$\{q\le0\}$ once it moves off $-a$), local stability of that branch
persists as long as its associated force stays within the elastic
interval, i.e.\ until $q_L(\ell)$ reaches the corner $q=0$; this happens
at $\ell=ka+\rho$, strictly larger than the energetic threshold
$\ell=\rho$ whenever $a>0$. A locally (balanced-viscosity) selected
solution would therefore remain on the left branch up to
$\ell=ka+\rho$ before jumping. Identifying the corresponding
BV post-jump state requires the vanishing-viscosity construction of
\cref{sec:bv} and is not carried out numerically here; the incremental
scheme~\eqref{eq:Ikstar}, minimising globally, realises instead the
energetic threshold $\ell=\rho$ of \cref{prop:jump_energetic}.
\end{remark}
\clearpage
