% =====================================================================
\section{Introduction}\label{sec:intro}
% =====================================================================

Epigenetic changes are heritable modifications of gene expression that
do not alter the DNA sequence, mediated for instance by DNA
methylation and histone modification~\cite{Bird2002}. Two features make
them distinctive from a modelling standpoint. First, they are
\emph{bistable}: cells commit to discrete phenotypes separated by
barriers, a picture formalised by Waddington's epigenetic
landscape~\cite{Waddington1957} and, in dynamical-systems terms, by
bistable regulatory switches~\cite{Huang2007,Ferrell2012}. Second, they
exhibit \emph{memory and hysteresis}: an epigenetic mark established in
response to a micro-environmental stimulus can persist after the
stimulus is removed, and its erasure requires crossing a distinct
threshold---as when hypoxia impairs the oxygen-dependent demethylation
machinery and leaves a durable methylation
imprint~\cite{Thienpont2016}. Threshold activation, path dependence and
residual state upon unloading are precisely the signatures of
\emph{rate-independent} dissipative evolution.

This observation is the starting point of the present work. We propose
to model epigenetic change within the mathematical theory of
rate-independent systems, a framework we call \emph{Rate-Independent
Epigenetics} (RIE). Rate-independent systems describe processes whose
response is invariant under time reparametrisation, governed by the
balance between a stored energy and a dissipation
mechanism~\cite{MielkeTheil,Mielke2005,MielkeRoubicek}. They were
developed largely in mechanics---plasticity~\cite{lubliner:1990,
hanreddy2013}, viscoelasticity~\cite{tschoegl1989},
damage~\cite{lemaitre2005uh}, fracture~\cite{francfort1998,dalmaso2006},
diffusion in porous media~\cite{bear1972}, frictional
contact~\cite{wriggers2006}, phase
transformation~\cite{balljames1987,mielke2002,carstensen2002uh} and
ferromagnetism~\cite{desimone2002,brokate1996}---and possess a
variational structure that yields powerful existence, uniqueness and
approximation results, valid even in the non-smooth
setting~\cite{MielkeRoubicek,visintin1994,Krasnoselskii}. The central
premise of RIE is that this same structure is the natural language for
epigenetic evolution: the double-well landscape of Waddington becomes
the stored energy, the micro-environment becomes the loading, and the
resistance of the epigenetic machinery to remodelling becomes the
dissipation threshold.

The appeal of this transfer is methodological as much as conceptual.
Once the three ingredients are fixed and an energetic evolution
principle is postulated, \emph{all} governing equations follow
systematically and are, by construction, thermodynamically consistent:
the first and second laws are not imposed a posteriori but are built
into the variational structure. This removes a recurrent difficulty in
quantitative epigenetics---the ad-hoc prescription of evolution
laws---and guarantees that any RIE model respects energy conservation
and non-negative entropy production irrespective of the choice of
landscape or dissipation.

Rate independence is, of course, an idealisation. We adopt it as the
\emph{quasi-static limit} of epigenetic evolution---the analogue of
perfect plasticity in mechanics---questionable in general, but a good
leading-order description whenever the chromatin state relaxes quickly
relative to its environment or stays frozen except at thresholds, and,
crucially, one that captures the threshold--memory--hysteresis
phenomenology while yielding a robust, thermodynamically consistent and
predictive skeleton. Rate-dependent effects, such as relaxation and
ratcheting, are recovered as a controlled viscous extension whose
vanishing-viscosity limit is the present model; we return to the
validity of the hypothesis, and to its relation with existing
paradigms, in \cref{sec:discussion}.

\paragraph{Scope and contributions.}
This paper develops the general RIE framework and its numerical
approximation; the calibrated, bistable biological model is the subject
of a companion paper. A model is a triple $(\Q,E,\Psi)$: a state space
$\Q\subseteq\R^n$ of epigenetic configurations, a stored-energy
functional $E$ depending on the state and on a loading path $S(t)$, and
a convex, $1$-homogeneous dissipation potential $\Psi$. Our
contributions are:
\begin{enumerate}[leftmargin=2em,itemsep=2pt]
  \item We derive the governing equations from the energetic principle:
  the energetic formulation (\cref{def:energetic}) and the equivalent
  Biot inclusion (\cref{prop:biot}), and give their thermodynamic
  reading---first law, second law, and a principle of minimal
  dissipation (\cref{sec:thermo}).
  \item We establish the well-posedness of RIE models
  (\cref{sec:math}): existence of energetic solutions, uniqueness under
  convexity, and existence and scalar uniqueness of balanced-viscosity
  solutions, which select the physically correct path at epigenetic
  switches.
  \item We construct and analyse a variational time integrator
  (\cref{sec:numerical}): convergence to energetic solutions and a
  global energy-consistency estimate, with the scalar update reducing
  to a two-sided return map.
  \item We validate the method on the linear play operator---the
  minimal RIE model of epigenetic memory---for which we give the exact
  closed-form evolution and reproduce both reversible and irreversible
  responses (\cref{sec:experiments}).
\end{enumerate}
To keep the narrative focused, the body states only the principal
results; supporting results and all proofs, in full detail, are
collected in the appendices.

\paragraph{Outline.}
\Cref{sec:framework} sets up the framework and derives the governing
equations. \Cref{sec:thermo} develops the thermodynamic interpretation.
\Cref{sec:math} presents the well-posedness theory and
\cref{sec:numerical} the numerical method and its analysis.
\Cref{sec:experiments} reports the numerical experiments;
\cref{sec:discussion} discusses the validity of the rate-independent
hypothesis, situates the framework among existing paradigms and
outlines its extensions; and \cref{sec:conclusions} concludes.
